---
title: "EDA"
params:
  path: ""
execute:
  warning: false
  message: false
format:
  html:
    theme: flatly
  pdf:
    fontsize: 11pt
df-print: kable
fig-width: 6
fig-height: 6
---

```{r}
#| echo: false

# Helper function to display a plot in Plotly if the report is in HTML format.
render_plot <- function(p) {
    if (knitr::is_html_output()) {
        plotly::ggplotly(p)
    } else {
        p
    }
}
```

```{r}
transaction <- arrow::read_parquet(file.path(params$path, "transaction.parquet"))
store <- arrow::read_parquet(file.path(params$path, "store.parquet"))
product <- arrow::read_parquet(file.path(params$path, "product.parquet"))
```

# Summaries
```{r}
describe <- skimr::skim_with(
    numeric = skimr::sfl(
        mean = ~ round(mean(., na.rm = TRUE), 2),
        sd = ~ round(sd(., na.rm = TRUE), 2),
        min = ~ round(min(., na.rm = TRUE), 2),
        max = ~ round(max(., na.rm = TRUE), 2),
        p0 = NULL, p50 = NULL, p100 = NULL, hist = NULL
    ),
    append = TRUE
)
```


## Transactions
```{r}
describe(tibble::as_tibble(transaction) |> dplyr::select(-week))
```


## Stores
```{r}
describe(store)
```


## Products
```{r}
describe(product)
```

# Outliers

```{r}
#' Outliers
#'
#' Print total number of outliers in transaction.
#' @param transaction A `tsibble` containing the transaction dataset.
#' @returns A `tibble` containing the outliers.
outliers <- function(transaction) {
    is_outlier <- function(x) {
        return(
            (x < quantile(x, 0.25, na.rm = TRUE) - 3 * IQR(x, na.rm = TRUE)) |
                (x > quantile(x, 0.75, na.rm = TRUE) + 3 * IQR(x, na.rm = TRUE))
        )
    }

    outliers <- transaction |>
        tibble::as_tibble() |>
        dplyr::reframe(
            across(
                .cols = c(units, visits, hhs, spend, price, base_price),
                .fns = is_outlier,
                .names = "outlier_{.col}"
            )
        ) |>
        dplyr::mutate(
            week = transaction$week,
            store_id = transaction$store_id,
            upc_id = transaction$upc_id,
            .before = 1
        )

    # Number of outliers per feature
    return(
        outliers |>
            dplyr::select(tidyselect::contains("outlier")) |>
            dplyr::reframe(dplyr::across(tidyselect::everything(), sum))
    )
}

outliers(transaction)
```

# What is the range of prices offered on product?

```{r}
#' Price range
#'
#' Print the range of prices offered on product.
#' @param transaction A `tsibble` containing the transaction data.
#' @returns A `skim` summary.
prices_range <- function(transaction) {
    # Skim summary custom function
    price_summary <- skimr::skim_with(
        numeric = skimr::sfl(
            min = ~ min(., na.rm = TRUE),
            max = ~ max(., na.rm = TRUE),
            mean = ~ mean(., na.rm = TRUE),
            sd = ~ sd(., na.rm = TRUE)
        ),
        append = FALSE
    )

    prices_range <- transaction |>
        tibble::as_tibble() |>
        dplyr::select(upc_id, price) |>
        dplyr::group_by(upc_id) |>
        price_summary()
    return(prices_range)
}

prices_range(transaction)
```

# What is the impact on sales of promotions, displays, or being featured in the circular?

```{r}
#' Product overview
#'
#' Helper functions to make visualizations
#' @param ts A `tibble` containing the transaction data.
#' @param selected_cat A `character` containing the selected product category.
#' @returns A `ggplotly` object.
product_overview <- function(ts, selected_cat) {
    return(
        ts |>
            dplyr::filter(category == selected_cat) |>
            tidyr::pivot_longer(
                c(units, base_price, feature, display, tpr_only),
                names_to = "col",
                values_to = "value"
            ) |>
            ggplot2::ggplot(ggplot2::aes(x = as.Date(tsibble::yearweek(week)), y = value, color = col)) +
            ggplot2::geom_line() +
            ggplot2::facet_wrap(~col, scales = "free_y", ncol = 1) +
            ggplot2::scale_fill_brewer(palette = "Set1") +
            ggplot2::labs(
                title = glue::glue("Product: {selected_cat}"),
                x = "Week"
            ) +
            ggplot2::theme(
                legend.position = "none",
                axis.text.x = ggplot2::element_text(angle = 45, hjust = 1)
            )
    )
}

#' Product correlation
#'
#' Plot a pairplot to see if there is any correlation among the features.
#' @param ts A `tibble` containing the transaction data.
#' @param selected_cat A `character` containing the selected product category.
#' @returns A `ggpairs` object.
product_corr <- function(ts, selected_cat) {
    return(
        ts |>
            dplyr::filter(category == selected_cat) |>
            dplyr::mutate(
                feature = ifelse(feature == 1, "True", "False"),
                display = ifelse(display == 1, "True", "False"),
                tpr_only = ifelse(tpr_only == 1, "True", "False")
            ) |>
            dplyr::select(
                -category,
                -week
            ) |>
            GGally::ggpairs() +
            ggplot2::labs(title = glue::glue("Product: {selected_cat}")) +
            ggplot2::theme_minimal(base_size = 9)
    )
}

# I will aggregate the data by product category, in order to
# compare promotions, displays and being featured against sales.
# Aggregation will lost information but the idea is to obtain an overview.
ts <- transaction |>
    dplyr::left_join(product, by = dplyr::join_by(upc_id)) |>
    tibble::as_tibble() |>
    dplyr::group_by(week, category) |>
    dplyr::summarise(
        .groups = "drop",
        units = sum(units, na.rm = TRUE),
        visits = sum(visits, na.rm = TRUE),
        base_price = mean(base_price, na.rm = TRUE),
        feature = ceiling(mean(feature, na.rm = TRUE)),
        display = ceiling(mean(display, na.rm = TRUE)),
        tpr_only = ceiling(mean(tpr_only, na.rm = TRUE))
    )
```

{{< pagebreak >}}

```{r}
product_overview(ts, "BAG SNACKS") |> render_plot()
```

{{< pagebreak >}}

```{r}
product_corr(ts, "BAG SNACKS") |> render_plot()
```

{{< pagebreak >}}

```{r}
product_overview(ts, "ORAL HYGIENE PRODUCTS") |> render_plot()
```

{{< pagebreak >}}

```{r}
product_corr(ts, "ORAL HYGIENE PRODUCTS") |> render_plot()
```

{{< pagebreak >}}

```{r}
product_overview(ts, "COLD CEREAL") |> render_plot()
```

{{< pagebreak >}}

```{r}
product_corr(ts, "COLD CEREAL") |> render_plot()
```

{{< pagebreak >}}

```{r}
product_overview(ts, "FROZEN PIZZA") |> render_plot()
```

{{< pagebreak >}}

```{r}
product_corr(ts, "FROZEN PIZZA") |> render_plot()
```

Observing the plots, we can see that exist correlation between units and promotions.
This is expected, since promotions are done to increase sales. But, some promoted product
don't show increase in sales. Further analysis is recommended for those product.
Also, is visible that units is slightly correlated with base_price, since
base_price represents the baseline price of the final price of the product.

{{< pagebreak >}}

# What is the impact on units/visits of promotions?

```{r}
#' Units/Visits promotion
#'
#' Helper functions to make visualizations.
#' @param ts A `tibble` of the transaction data.
#' @param selected_cat A `character` containing the product category to be selected.
#' @returns A `ggplotly` object.
units_visits_prom <- function(ts, selected_cat) {
    return(
        ts |>
            dplyr::filter(category == selected_cat) |>
            dplyr::mutate(units_visits = units / visits) |>
            tidyr::pivot_longer(
                c(units_visits, feature, display, tpr_only),
                names_to = "col",
                values_to = "value"
            ) |>
            ggplot2::ggplot(ggplot2::aes(x = as.Date(tsibble::yearweek(week)), y = value, color = col)) +
            ggplot2::geom_line() +
            ggplot2::facet_wrap(~col, scales = "free_y", ncol = 1) +
            ggplot2::scale_fill_brewer(palette = "Set1") +
            ggplot2::labs(
                title = glue::glue("Product: {selected_cat}"),
                x = "Week"
            ) +
            ggplot2::theme(
                legend.position = "none",
                axis.text.x = ggplot2::element_text(angle = 45, hjust = 1)
            )
    )
}
```

{{< pagebreak >}}

```{r}
units_visits_prom(ts, "BAG SNACKS") |> render_plot()
```

{{< pagebreak >}}

```{r}
units_visits_prom(ts, "ORAL HYGIENE PRODUCTS") |> render_plot()
```

{{< pagebreak >}}

```{r}
units_visits_prom(ts, "COLD CEREAL") |> render_plot()
```

{{< pagebreak >}}

```{r}
units_visits_prom(ts, "FROZEN PIZZA") |> render_plot()
```

In aggregated TS there isn't visible too much correlation among the
units/visits and promotions because aggregated product are very noisy
in promotional features.
Further analysis along each TS individually (disaggregated) is recommended.
